% -------------------------------------------------------- %
% UDEQueue Lock
% by: Isai Barajas Cicourel


% -------------------------------------------------------- %
% Document Start

\section{\textbf{Unbounded Double Ended Queue}}


% -------------------------------------------------------- %
% Particular Caes

\subsection{Particular Case}
\par
In this experiment we implement an unbounded double queue based on the previous double queue implemented.
\par


% -------------------------------------------------------- %
% Solution Information

\subsection{Solution}
\par
According to the theory the \textit{UnboundedDEQueue} class dynamically resizes itself as needed.
We implement the \textit{UnboundedDEQueue} in a cyclic array, with top and bottom fields as in the \textit{BoundedDEQueue} (except indexed modulo the array's capacity). As before, if bottom is less than or equal to top, the \textit{UnboundedDEQueue} is empty.
The difference resides in the \textit{CircularArray()} class which provides \textit{get()} and \textit{put()} methods that add and remove tasks, and a \textit{resize()} method that allocates a new circular array and copies the old array's contents into the new array.
\par
\begin{lstlisting}[frame=single,breaklines=true]
  class CircularArray {
    private int logCapacity;
    private Runnable[] currentTasks;
    CircularArray(int logCapacity) {
      this.logCapacity = logCapacity;
      currentTasks = new Runnable[1 << logCapacity];
    }
    int capacity() {
      return 1 << this.logCapacity;
    }
    Runnable get(int i) {
      return this.currentTasks[i % capacity()];
    }
    void put(int i, Runnable task) {
      this.currentTasks[i % capacity()] = task;
    }
    CircularArray resize(int bottom, int top) {
      CircularArray newTasks =
          new CircularArray(this.logCapacity+1);
      for (int i = top; i < bottom; i++) {
        newTasks.put(i, this.get(i));
      }
      return newTasks;
    }
  }
\end{lstlisting}

% -------------------------------------------------------- %
% Experiment

\subsection{Experiment Description} 
\par
The test creates $16$ threads, eight that need to be coordinate in order to \textit{pushBottom()} and \textit{popBottom()} an array of values. All threads have to cooperate to add and remove elements from the queue and eight \textit{Dummy()} threads will be stealing execution from the running ones. After each thread execution, if everything works according to the test the map will have all values set to true.
If that is not the case, a duplicate or missing fail will be raised.
\par


% -------------------------------------------------------- %
% Results

\subsection{Observations and Interpretations}

\par
The test needs some changes in order to compile, once this changes are implemented the test executes as expected.
\begin{lstlisting}[frame=single,breaklines=true]
Tests run: 2, Failures: 0, Errors: 0, Skipped: 0, Time elapsed: 1.4 sec

------------- Standard Output ---------------
sequential pushBottom and popBottom
concurrent pushBottom and popBottom
------------- ---------------- ---------------
test-single:
BUILD SUCCESSFUL (total time: 1 second)
\end{lstlisting}


% -------------------------------------------------------- %
% Results

\subsection{Proposed changes to fix the problem}

\par
The UDEQueueTest class needs to have replaced the calls to non defined reset by new instance in order to compile.
\begin{lstlisting}[frame=single,breaklines=true]
  //instance.reset();
  instance = new UDEQueue(THREADS);
\end{lstlisting}

