% --------------------------- %
% FineGrainedHeapTest Start
% --------------------------- %
\section{\textbf{FineGrainedHeapTest}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Particular Case}
\par
Now we look at the problem of implementing an Unbounded Priority Queue in a
Fine-Grained fashion. 
\par
Again, remember that Unbounded in this context means that the range is
unbounded, for example the set of 32-bits integers or floating point values.
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solution}
\par
The solution proposed in this exercise is the Fine-Grained heap. The particular
characteristic of this implementation is that percolation happens in a sequence
of discrete atomic steps that can be interleaved with other such steps. The same
happens with the \textit{removeMin()} method.
\par
The \textit{add()} method first locks the heap and then takes the lock to the
next available child and puts the new value there. After that, it releases the
heap lock. After this point, it must percolate this new node. To do this, it has
to swap child and parent after locking them. This only happens if the parent is
available and the child has a higher priority. Otherwise, the node is already
where it belongs. At the end of this step, child and parent are unlocked. Then
it goes up one node in the heap, and does the same.
\par
\hfill
\begin{lstlisting}[style=numbers]
  /**
   * Add item to heap.
   * @param item Uninterpreted item.
   * @param priority item priority
   */
  public void add(T item, int priority) {
    heapLock.lock();
    int child = next++;
    heap[child].lock();
    heapLock.unlock();
    heap[child].init(item, priority);
    heap[child].unlock();
    while (child > ROOT) {
      int parent = child / 2;
      heap[parent].lock();
      heap[child].lock();
      int oldChild = child;
      try {
        if (heap[parent].tag == Status.AVAILABLE && heap[child].amOwner()) {
          if (heap[child].score < heap[parent].score) {
            swap(child, parent);
            child = parent;
          } else {
            heap[child].tag = Status.AVAILABLE;
            heap[child].owner = NO_ONE;
            return;
          }
        } else if (!heap[child].amOwner()) {
          child = parent;
        }
      } finally {
        heap[oldChild].unlock();
        heap[parent].unlock();
      }
    }
    if (child == ROOT) {
      heap[ROOT].lock();
      if (heap[ROOT].amOwner()) {
        heap[ROOT].tag = Status.AVAILABLE;
        heap[child].owner = NO_ONE;
      }
      heap[ROOT].unlock();
    }
  }

\end{lstlisting}
\hfill
\par
The \textit{removeMin()} method first locks the heap, the bottom and the root
nodes. If the root is empty, then we are done. Otherwise, it swaps root and
bottom and unlocks bottom. After that, it has to percolate the new root
element. It follows a similar approach as the one mentioned in the
\textit{add()} method. One difference is that this percolation goes from top to
bottom.
\par
\hfill
\begin{lstlisting}[style=numbers]
  /**
   * Returns and removes lowest-priority item in heap.
   * @return lowest-priority item.
   */
  public T removeMin() {
    heapLock.lock();
    int bottom = --next;
    heap[bottom].lock();
    heap[ROOT].lock();
    heapLock.unlock();
    if (heap[ROOT].tag == Status.EMPTY) {
      heap[ROOT].unlock();
      heap[bottom].lock();
      return null;
    }
    T item = heap[ROOT].item;
    heap[ROOT].tag = Status.EMPTY;
    swap(bottom, ROOT);
    heap[bottom].owner = NO_ONE;
    heap[bottom].unlock();
    if (heap[ROOT].tag == Status.EMPTY) {
      heap[ROOT].unlock();
      return item;
    }
    int child = 0;
    int parent = ROOT;
    while (parent < heap.length / 2) {
      int left = parent * 2;
      int right = (parent * 2) + 1;
      heap[left].lock();
      heap[right].lock();
      if (heap[left].tag == Status.EMPTY) {
        heap[right].unlock();
        heap[left].unlock();
        break;
      } else if (heap[right].tag == Status.EMPTY || heap[left].score < heap[right].score) {
        heap[right].unlock();
        child = left;
      } else {
        heap[left].unlock();
        child = right;
      }
      if (heap[child].score < heap[parent].score) {
        swap(parent, child);
        heap[parent].unlock();
        parent = child;
      } else {
        heap[child].unlock();
        break;
      }
    }
    heap[parent].unlock();
    return item;
  }
\end{lstlisting}
\hfill
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment Description}
\par
The same three test cases were provided in this experiment, so let us go
directly to the results.
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sample Results and Interpretation}
\par
Here is a sample output of the test demonstrating that the implementation of
this concurrent heap works properly:
\par
\hfill
\begin{verbatim}
[oraadm@gdlaa008 ch15]$ junit priority.FineGrainedHeapTest
.testSequential
OK.
.testParallelBoth
OK.
.testParallelAdd
OK.

Time: 3.175

OK (3 tests)
\end{verbatim}
\hfill
\par
One interesting thing about this exercise is that it takes longer than other
tests. Actually, other students shared the results with us and all of them
experimented different execution times. For example:
\par
\begin{itemize}
\item Our machine, that has 24cores, executed the test in $3.175$ seconds.
\item On a 6-cores machine, it took 1-6 secs.
\item On a 2-cores machine, it took 8-9 secs.
\item On a 100-cores machine it took $> 5$min and it did not finish.
\end{itemize}
\par
As a future work, we might want to understand why we had these execution times.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --------------------------- %
% FineGrainedHeapTest End
% --------------------------- %
