% --------------------------- %
% StripedHashSetTest Start
% --------------------------- %
\section{\textbf{StripedHashSetTest Test}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Particular Case}
\par
In this exercise we deal with the problem of building a hash data structure that
can be accessed by multiple threads without losing performance. 
\par
In the coarse version of our hash, we saw that any access to the data structure
ended up being serial. So, in other words, in this experiment we want to
understand how to create a hash that allows more concurrent accesses.
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solution}
\par
A solution for this problem is the \textit{Striped Hash}. The main
characteristic of this implementation is that the data structure contains an
array of locks. Each lock protect specific sections of the hash table. Here is
the implementation of the \textit{acquire()} and \textit{release()} methods:
\par
\hfill
\begin{lstlisting}[style=numbers]
  /**
   * Synchronize before adding, removing, or testing for item
   * @param x item involved
   */
  public final void acquire(T x) {
    int myBucket = Math.abs(x.hashCode() % locks.length);
    locks[myBucket].lock();
  }

  /**
   * synchronize after adding, removing, or testing for item
   * @param x item involved
   */
  public void release(T x) {
    int myBucket = Math.abs(x.hashCode() % locks.length);
    locks[myBucket].unlock();
  }
\end{lstlisting}
\hfill
\par
Note that when we acquire or release a lock, first we need to choose which lock
we should work with. 
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment Description}
\par
The proposed experiment is exactly the same as the one described in the previous
experiment. As mentioned before, there are four variants:
\textit{testSequential}, \textit{testParallelEnq}, \textit{testParallelRemove}
and \textit{testParallelBoth}.
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sample Results and Interpretation}
\par
Here is the result that we got:
\hfill
\begin{verbatim}
[oraadm@gdlaa008 Hash]$ junit hash.StripedHashSetTest
.sequential add, contains, and remove
.parallel both
.parallel add
.parallel remove

Time: 0.027

OK (4 tests)
\end{verbatim}
\hfill
\par
However, some people in the group mentioned that they had the same test issue as
mentioned in the Coarse Hash Set exercise. Again, we think that it is a test
issue since there is no garantee that removers will remove elements that have
been really inserted.
\par
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --------------------------- %
% StripedHashSetTest End
% --------------------------- %
